---
title: "Chemotaxis - two strains"
output:
word_document: default
html_document:
df_print: paged
pdf_document: default
---

## Chemotaxis Assay - Two strains


This week we will collect data for chemotaxis behavior of wildtype v daf-19(-) C. elegans to 1% isoamyl alcohol. The goals for this week are to:
*calculate a chemotaxis index for each plate
*calculate a mean chemotaxis index for each strain with 95% confidence intervals
*create a bar graph with this data
*perform a Welch's t-test to determine if the mean chemotaxis index for these strains are significantly different

Load the tidyverse:
```{r}
library(tidyverse)

```


First you will need to upload your data for both strains in a .csv file. Upload load your data file in the lower right window using the upload button, and then load the data using the code below in a data frame called allData. Your file name should be in the "".



```{r}

#allData<-read_csv("data.csv")

#allData

```


First, we want to determine the Chemotaxis Index for each plate. 
The formula is:

ChemI= ((T1+T2)-(C1+C2))/(T1+T2+C1+C2)



We can add a column to our data frame by using the "mutate" function. "Mutate" allows you to create a new column of values based upon data already in the data frame. So using our example above:



```{r}


#df<-mutate(.data=allData, ChemI=((T1+T2)-(C1+C2))/(T1+T2+C1+C2))

#print out your updated df dataframe
#df

```


Now that we have the Chemotaxis Index for each plate, we can use group by/summarise to calculate the average, standard deviation, n, and upper and lower confidence intervals:

### Calculate means for the different strains.

```{r}
#meanChemI<-df %>% group_by(strain) %>% summarise(avg=mean(ChemI), StandardDeviation=sd(ChemI), n=n(), upperCI=avg+((qnorm(0.975)*StandardDeviation)/sqrt(n)), lowerCI=avg-((qnorm(0.975)*StandardDeviation)/sqrt(n)))

#meanChemI


```
### Now make a barplot using ggplot
```{r}
#ggplot(data=meanChemI, mapping=aes(x=strain, y=avg))+geom_bar(stat="identity")


```
# Take a look at your plot.  Is this the order you want the data presented?  The code has plotted the data in the order it appeared in the meanChemI data table, daf-19 followed by wt.  Generally, the data for the wild-type strain should appear first on the x-axis followed by any mutant strains. To make this change, use the following code to determine the order of the x-axis and then make the plot again.

```{r}
#meanChemI$strain<-factor(meanChemI$strain, levels=c("wt", "daf-19"))

#ggplot(data=meanChemI, mapping=aes(x=strain, y=avg))+geom_bar(stat="identity")



```


# Improve the bargraph with labels for the axes.  You do not need a title for your graph in R.  For the chemotaxis project, your final format for presenting your data will be an oral presentation with a slideshow. You will add a title later when you generate Slides to present your results.

#The theme_minimal code is a choice for how the graph background and axes look. If you want to play with how the graph looks, you can do a websearch for geom_bar themes and locate other code options to designate a different presentation of the data.
```{r}
#ggplot(data=meanChemI, mapping=aes(x=strain, y=avg))+geom_bar(stat="identity")+labs(x="your x-axis label", y="your y-axis label", fill = "Strain") + theme_minimal()+ylim(0,1)

```



### Plot a bar graph with 95% Confidence interval used for error bars and adjust the y-axis limits to extend from -1 to +1.

```{r}
#ChemIPlot <- ggplot(data=meanChemI, aes(strain, avg)) +
#geom_col() +
#geom_errorbar(aes(ymin = lowerCI, ymax = upperCI), width=0.2)

#ChemIPlot + labs(x="your x-axis label", y="your y-axis label", fill = "Strain") + theme_minimal() + ylim(-1,1)


```
### Save the plot
```{r}
ggsave("plot.png", height=5, width=8)


```
###Look at your plot. You have two bars representing the mean chemotaxis index for each strain. How do you know if these are statistically different from each other? 

#from the Handbook for Biological Statistics by John H. MacDonald (http://www.biostathandbook.com/probability.html):

#"The basic idea of a statistical test is to identify a null hypothesis, collect some data, then estimate the probability of getting the observed data if the null hypothesis were true. If the probability of getting a result like the observed one is low under the null hypothesis, you conclude that the null hypothesis is probably not true."

#The null hypothesis for comparing two samples is that they are not different. You must run a statistical analysis called a t-test to calculate the probability (p-value) that the null hypothesis is true, i.e., the strains have the same chemotaxis index. A very low p-value indicates that the strains have very low probability of being the same and are instead significantly different. A Welch's t-test is specifically used when sample sizes are small (less than 10) or unequal or if variances are large.

#t.test(ChemI ~ strain, data=df) will examine the chemotaxis index, "ChemI", of each plate sorted by "~" the strain in our df dataframe table created earlier in this tutorial to perform a Welch's t-test with your data.

#Generally a p-value less than 0.05 indicates the two means are significantly different and in this case that your mutant strain behaves differently than your wild-type strain.
```{r}
#t.test(ChemI ~ strain, data=df)

```
# Be sure to record the results of your t-test in your online lab notebook.
