---
title: "Chemotaxis - two strains"
output:
word_document: default
html_document:
df_print: paged
pdf_document: default
---

## Chemotaxis - Two strains


You have been given data for chemotaxis behavior of wildtype v daf-19 C. elegans to isoamyl alcohol.  You will use this data to calculate a chemotaxis index for each plate.  Then you will calculate a mean chemotaxis index for each strain with 95% confidence intervals.  Finally, you will perform a Welch's t-test to determine if the mean chemotaxis index for isoamyl alcohol for these strains are significantly different.

Load the tidyverse:
```{r}
library(tidyverse)

```


For this lab, the the collected data for both strains is available in the data.csv file. Below, load the data, and then look at the data frame.



```{r}

allData<-read_csv("data.csv")

allData

```


Like last lab, we want to determine the Chemotaxis Index. 
The formula is:

ChemI= ((T1+T2)-(C1+C2))/(T1+T2+C1+C2)

Since our data is all in one data frame and contains more than one strain, we have to do a little more work to get our 


We can add a column to our data frame by using the "mutate" function. "Mutate" allows you to create a new column of values based upon data already in the data frame. So using our example above:



```{r}


df<-mutate(.data=allData, ChemI=((T1+T2)-(C1+C2))/(T1+T2+C1+C2))

#print out our updated df dataframe
df


```


Now that we have the Chemotaxis Index for each plate, we can use group by/summarise to calculate the average, standard deviation, n, and upper and lower confidence intervals:

### Calculate means for the different strains.

```{r}
meanChemI<-df %>% group_by(strain) %>% summarise(avg=mean(ChemI), StandardDeviation=sd(ChemI), n=n(), upperCI=avg+((qnorm(0.975)*StandardDeviation)/sqrt(n)), lowerCI=avg-((qnorm(0.975)*StandardDeviation)/sqrt(n)))

meanChemI
```
### Now make a barplot using ggplot
```{r}
ggplot(data=meanChemI, mapping=aes(x=strain, y=avg))+geom_bar(stat="identity")
```
# Improve the bargraph with title and labels and pick a theme
```{r}
ggplot(data=meanChemI, mapping=aes(x=strain, y=avg))+geom_bar(stat="identity")+labs(title="Chemotaxis index of C. elegans strains to 1% Isoamyl alcohol after 1 hour", x="C elegans strain", y="Chemotaxis Index", fill = "Strain") + theme_minimal()+ylim(0,1)
```



### Plot a bar graph with 95% Confidence interval used for error bars

```{r}
ChemIPlot <- ggplot(data=meanChemI, aes(strain, avg)) +
geom_col() +
geom_errorbar(aes(ymin = lowerCI, ymax = upperCI), width=0.2)

ChemIPlot + labs(title="Chemotaxis index of C. elegans strains to 1% Isoamyl alcohol after 1 hour",y="Chemotaxis Index (95% CI)", x = "Strain") + theme_minimal() + ylim(0,1)
```
### Save the plot
```{r}
ggsave("plot.png", height=5, width=8)


```

Now run a Welch two sample t-test to determine whether the measured CI values are significantly different:

```{r}
t.test(ChemI ~ strain, data=df)

```




